Map{R}educe in {MPI} for Large-scale Graph Algorithms

### 2010-04-15 paper/graph.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@549 3fc257bb-8550-0410-81b3-a5c103da83f7

@@

@@

@@

 \caption{MapReduce algorithm for connected component labelling.}

@@

The first {\it map()} operation emits the vertices in each edge as
keys, with the edge as a value.  The current zone assignment of each
vertex is added to the set of key/value pairs.  The first collate()
operation collects all the edges of a vertex and its zone assignment
together in one multi-value.  The first reduce operation then re-emits
each edge, tagged by the zone assignment of one of its vertices.

Since each edge was emitted twice, the second collate operation
collects the zone assignments for its two vertices together.  If the
two zone IDs are different, the second reduce operation chooses a
winner (the min of the 2 IDs), and emits the loser ID as a key, with
the winning ID as a value.  If no zone ID changes are emitted, the
algorithm is finished, and the iteration exits.

The third map() operation inverts the vertex/zone key/value pairs to
become zone/vertex pairs.  The third add operation adds the changing
zone assignments to the set of key/value pairs.  The third collate()
can then collects all the vertices of a zone and zero or more
reassignments for the zone ID.  Since a zone could collide with
multiple other zones on the same iteration due to shared edges, the
new zone ID becomes the minimum ID of any of the neighboring zones.
If no zone reassignment value appears in the multi-value, the zone ID
is unchanged.  The final reduce() a key/value pair for each vertex in
the zone, with the vertex as a key and the new zone ID as a value.

Note that if a graph has only a few components, then the third collate
operation, which keys on the zone ID, may generate a few very large
key/multi-value (KMV) pairs.  For example, if the graph is fully
connected, then on the last iteration, a single KMV pair will contain
all vertices in the graph and be assigned to one processor.  This
imbalance in memory and computational work can lead to poor parallel
performance of the overall algorithm.  To counter this effect, the
various operations of stage 3 include extra logic.  The idea is to
partition zones whose vertex count exceeds a user-defined threshhold
into $P$ sub-zones, where $P$ is the number of processors.  The 64-bit
integer that stores the zone ID also stores a bit flag indicating the
zone has been partitioned and a set of bits that encode the processor
ID.

During the third map() operation, if the zone has been partitioned,
then the vertex is assigned to a random processor and the processor ID
bits are added to the zone ID, as indicated by the $Z_i^+$ notation in
Figure \ref{fig:cc}.  Likewise, if $Z_i$ has been partitioned, the
third add() operation emits the zone ID change key/value pair
($Z_i$,$Z_{winner}$) as ($Z_i^+$,$Z_{winner}$).  In this case
($Z_i$,$Z_{winner}$), is emitted not once, but $P+1$ times, once for
each processor, and once as if $Z_i$ had not been partitioned (for a
reason discussed below).

This additional partitioning logic means that the third collate()
operation, which keys on zone IDs, some of which now include processor
bits, will collect only a $1/P$ subset of the vertices in large zones
onto each processor.  But the multivalue on each processor contains
all the zone-reassignments relevant to the unpartitioned zone.  This
allows the third reduce operation to change the zone ID (if necessary)
in a consistent manner across all $P$ multi-values that contain the
zone's vertices.  When the reduce() operation emits new zone
assignments for each vertex, the zone retains its partitioned status,
and the partition bit is also explicitly set if the vertex count
exceeds the threshhold for the first time.

Note that this logic does not guarantee that the partition bits of the
zone IDs for all the vertices in a single zone will be set
consistently on a given iteration.  For example, an unpartitioned zone
with a small ID may consume a partitioned zone.  The vertices from the
partitioned zone will retain their partitioned status, but the
original vertices in the small zone may not set the partition bit of
their zone IDs.  On subsequent iterations, the third add() operation
emits $P+1$ copies of new zone reassignments for both partitioned and
unpartioned zone IDs, to insure all vertices in the zone will know the
reassignment information.

Discussion of Luby algorithm for maximally independent sets.

footnote on NP-complete for maximal; MapReduce is unlikely to
help with that issue.

@@

 \caption{MapReduce algorithm for R-MAT graog generation.}

@@

>>>>

\subsection{PageRank}

@@

\subsection{Triangle enumeration}

@@

\subsection{Connected component labeling}

@@

 \caption{MapReduce algorithm for connected component labeling.}

@@

The first {\it map()} operation emits the vertices in each edge as
keys, with the edge as a value.  The current zone assignment of each
vertex is added to the set of key/value pairs.  The first {\it
collate()} operation collects all the edges of a vertex and its zone
assignment together in one multi-value.  The first {\it reduce()}
operation then re-emits each edge, tagged by the zone assignment of
one of its vertices.

Since each edge was emitted twice, the second {\it collate()}
operation collects the zone assignments for its two vertices together.
If the two zone IDs are different, the second {\it reduce()} operation
chooses a winner (the min of the 2 IDs), and emits the loser ID as a
key, with the winning ID as a value.  If no zone ID changes are
emitted, the algorithm is finished, and the iteration exits.

The third {\it map()} operation inverts the vertex/zone key/value
pairs to become zone/vertex pairs.  The third {\it add()} operation
adds the changing zone assignments to the set of key/value pairs.  The
third collate() can then collects all the vertices of a zone and zero
or more reassignments for the zone ID.  Since a zone could collide
with multiple other zones on the same iteration due to shared edges,
the new zone ID becomes the minimum ID of any of the neighboring
zones.  If no zone reassignment value appears in the multi-value, the
zone ID is unchanged.  The final reduce() a key/value pair for each
vertex in the zone, with the vertex as a key and the new zone ID as a
value.

Note that if a graph has only a few components, then the third {\it
collate()} operation, which keys on the zone ID, may generate a few
very large key/multi-value (KMV) pairs.  For example, if the graph is
fully connected, then on the last iteration, a single KMV pair will
contain all vertices in the graph and be assigned to one processor.
This imbalance in memory and computational work can lead to poor
parallel performance of the overall algorithm.  To counter this
effect, the various operations of stage 3 include extra logic.  The
idea is to partition zones whose vertex count exceeds a user-defined
threshhold into $P$ sub-zones, where $P$ is the number of processors.
The 64-bit integer that stores the zone ID also stores a bit flag
indicating the zone has been partitioned and a set of bits that encode
the processor ID.

During the third {\it map()} operation, if the zone has been
partitioned, then the vertex is assigned to a random processor and the
processor ID bits are added to the zone ID, as indicated by the
$Z_i^+$ notation in Figure \ref{fig:cc}.  Likewise, if $Z_i$ has been
partitioned, the third {\it add()} operation emits the zone ID change
key/value pair ($Z_i$,$Z_{winner}$) as ($Z_i^+$,$Z_{winner}$).  In
this case ($Z_i$,$Z_{winner}$), is emitted not once, but $P+1$ times,
once for each processor, and once as if $Z_i$ had not been partitioned
(for a reason discussed below).

This additional partitioning logic means that the third {\it
collate()} operation, which keys on zone IDs, some of which now
include processor bits, will collect only a $1/P$ subset of the
vertices in large zones onto each processor.  But the multivalue on
each processor contains all the zone-reassignments relevant to the
unpartitioned zone.  This allows the third {\it reduce()} operation to
change the zone ID (if necessary) in a consistent manner across all
$P$ multi-values that contain the zone's vertices.  When the {\it
reduce()} operation emits new zone assignments for each vertex, the
zone retains its partitioned status, and the partition bit is also
explicitly set if the vertex count exceeds the threshhold for the
first time.

Note that this logic does not guarantee that the partition bits of the
zone IDs for all the vertices in a single zone will be set
consistently on a given iteration.  For example, an unpartitioned zone
with a small ID may consume a partitioned zone.  The vertices from the
partitioned zone will retain their partitioned status, but the
original vertices in the small zone may not set the partition bit of
their zone IDs.  On subsequent iterations, the third {\it add()}
operation emits $P+1$ copies of new zone reassignments for both
partitioned and unpartioned zone IDs, to insure all vertices in the
zone will know the reassignment information.

\subsection{Maximal independent set identification}

An ``independent'' set of vertices in a graph is one where no pair of
vertices in the set share an edge.  The set is ``maximal'' if no
additional vertex can be added to it \footnote{A maximal set is
``maximum'' if it is the largest maximal set.  Finding a graph's
maximum independent set is an NP-hard problem, which MapReduce is
unlikely to help with.}.  Finding a maximal independent set is useful
in several contexts, such as identifying large numbers of indepenent
starting vertices for graph traversals.  Luby's algorithm \cite{Luby}
is a well-known parallel method for finding a minimal set.  Figure
\ref{fig:luby} outlines a MapReduce version of Luby's algorithm.

@@

 \caption{MapReduce algorithm for finding amaximal independent set of
 graph vertices.}

 \label{fig:luby}
\end{figure}

@@

 \label{fig:sssp}
\end{figure}
